---
title: "Effect of age and light on the antioxidant activity of *Ilex guayusa* leaves - Statistical analysis"
author: "Jefferson Pastuña"
date: "2025-01-31"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

# Introduction

This R Script aims to record the procedure for analyzing the antioxidant activity of *Ilex guayusa* leaves under different age and light conditions. Each step has a brief explanation, as well as code and graphics.

The logistic model proposed by [Study of the DPPH-scavenging activity: Development of a free software for the correct interpretation of data](https://doi.org/10.1016/j.foodchem.2008.10.035) was used to interpret the antioxidant activity result of *Ilex guayusa* leaves. We do not use the logistic model to calculate the IC50. However, we use the logistic model (multivariable logistic model) to understand the relationship between age-light factors and antioxidant activity.

# Data preprocessing

## Linearity

Before the data analysis, we checked the linear relationship between the sample concentration and the antioxidant activity (measured as an inhibition percentage). We verified the linear relationship between the sample concentration and the antioxidant activity because we required verification that the sample concentration did not interfere with the independent variables (age and light factor). Also, the logistic regression assumes the linearity of independent variables and log odds of the dependent variable.

First, we loaded the data to R.

```{r data, echo=TRUE}

# Installing readxl package
#install.packages("readxl")
# Loading readxl package
library("readxl")
# Loading antioxidant data
abts <- read_excel("Data/Antioxidant_data/Antioxidant_to_R.xlsx",
                   sheet = "ABTS_data")
# Change character variables as factor variables
abts$Analysis_day <- as.factor(abts$Analysis_day)
abts$Location <- as.factor(abts$Location)
abts$Age <- as.factor(abts$Age)
abts$Light <- as.factor(abts$Light)
# Loading dplyr package
library(dplyr)
# Visual inspection of data
glimpse(abts)

```

Then, we use the glm function to fit generalized linear models.

```{r, echo=TRUE}

# Build the model
abts_concentration <- glm(formula = Inhibition_persentage ~ Concentration,
                          data = abts,
                          family = quasibinomial(logit))

```

Summary of the model.

```{r, echo=TRUE}

# Summary of the model
summary(abts_concentration)

```

The slope (m = 0.0059967) showed that the antioxidant activity increased with the sample concentration, and this increase is significant (p-value <2e-16). We plot the model using the following line code.

```{r, echo=TRUE}

# Installing sjPlot package
#install.packages("sjPlot")
# Loading sjPlot package
library(sjPlot)
# Plot of the model
concentration_plot <-
  plot_model(abts_concentration, type = "pred", terms = "Concentration",
             #show.data = TRUE, jitter = 0.01,
             title = "Predicted probabilities of inhibition persentage",
             axis.title = c("Concentration", "Inhibition persentage"))
concentration_plot

```

The next step will be to confirm the linearity between the sample concentration and the antioxidant activity.

```{r, echo=FALSE, results='asis'}

# Use polynomial degrees
abts_conc_pol2 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 2),
                      data = abts,
                      family = quasibinomial(logit))
abts_conc_pol3 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 3),
                      data = abts,
                      family = quasibinomial(logit))
abts_conc_pol4 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 4),
                      data = abts,
                      family = quasibinomial(logit))
# Display degrees
concentration_models <- tab_model(abts_conc_pol4)
concentration_models

```

The results showed that the sample concentration and the antioxidant activity could not have a linear relationship. The result showed that the linear (p-value <0.001) and polynomial grade 2 (p-value <0.001) models fit the data better.

## Collinearity

Before the data analysis, we also check the collinearity of the whole data. We use the glm function to fit generalized linear models with all variables (age, light, location, and analysis day). The analysis day (due to the extended samples, the analysis of antioxidant activity takes several days) was included to inspect the collinearity with other variables.

```{r, echo=TRUE}

# Build the model
abts_all <- glm(formula = Inhibition_persentage ~ Location + Age + Light +
                  Analysis_day,
                data = abts,
                family = quasibinomial(logit))

```

The following code helps us to check all model assumptions visually.

```{r, echo=TRUE, fig.width = 8, fig.height = 6}

# Installing performance package
#install.packages("performance")
# Loading performance package
library(performance)
# Check all model assumptions visually
abts_check <- check_model(abts_all)
abts_check

```

We can observe that all residuals are in the error bands. The influential observations plot showed the absence of outliers, the collinearity plot showed the colinear relationship between analysis day and location factor, and the "normality of residuals" plot showed that the data do not have a normal distribution. Due to the collinearity between the analysis day and the location factor, we exclude the location factor from the analysis.

# Data analysis

After data preprocessing, the age and light factors are the suitable variables for use in the multivariable logistic regression. Again, we use the glm function to fit generalized linear models with age and light factors.

```{r, echo=TRUE}

# Build the model
abts_m <- glm(formula = Inhibition_persentage ~ Age + Light,
              data = abts,
              family = quasibinomial(logit))

```

## Model assumptions

The following code helps us to check all model assumptions visually.

```{r, echo=TRUE, fig.width = 8, fig.height = 6}

# Check all model assumptions visually
abts_m_check <- check_model(abts_m)
abts_m_check

```

We can observe that all residuals fall within the error bands. The "influential observations" plot shows that the points remain inside the contour lines. The collinearity plot showed that all two variables are not multicollinear, as the variation inflation factor stays below 5. The "normality of residuals" plot could suggest the normal distribution of the data. Since our model's assumptions are satisfied, we can visualize the model results.

## Model results

The following code extracts the prediction (inhibition percentage) for the two study variables.

```{r, echo=TRUE}

# Installing ggeffects package
#install.packages("ggeffects")
# Loading ggeffects package
library(ggeffects)
# Extract prediction
abts_m_pred <- ggeffect(abts_m)
abts_m_pred

```

Plot the results.

```{r, echo=TRUE}

# Library to use "ggsave"
library(tidyverse)
# Visualize predictions
abts_plot <- ggeffect(abts_m) %>%
  plot(show_title = F, show_x_title = F)  %>%
  sjPlot::plot_grid()
# Safe result plot
ggsave('Result/Antioxidant_Result/ABTS_Result.pdf', plot = abts_plot,
       width = 8, height = 6, device='pdf', dpi="print")
ggsave('Result/Antioxidant_Result/ABTS_Result.png', plot = abts_plot,
       width = 8, height = 6, device='png', dpi="print")

```

Print the result in a table.

```{r, echo=TRUE}

# Installing gtsummary package
#install.packages("gtsummary")
# Loading gtsummary package
library(gtsummary)
# Display contrasts / post-hocs / pairwise camparisons
abts_m_table <- tbl_regression(abts_m, exponentiate = TRUE,
                               add_pairwise_contrasts = TRUE,
                               contrasts_adjust = "bonferroni",
                               pairwise_reverse = FALSE,
                               pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p()
abts_m_table

```

Print the equation of logistic regression.

```{r, echo=TRUE}

# Installing equatiomatic package
#install.packages("equatiomatic")
# Loading equatiomatic package
#library(equatiomatic)
# Produce model equation easily
extract_eq(abts_m)

```








```{r cars}


 check_normality(abts_m)

simulate_residuals(abts_m)

check_residuals(abts_m)

# Visual inspection of data
glimpse(abts)

# Build the model
abts_m <- glm(formula = Inhibition_persentage ~ Concentration, #Location_factor + Age_factor +
                #Light_factor + Concentration + Analysis_day,
              data = abts,
              family = binomial)

# Installing sjPlot package
#install.packages("sjPlot")

# Loading sjPlot package
library(sjPlot)

# Check non-linearity
summary(abts_m)

plot_model(abts_m, type = "pred", terms = "Concentration")

# Use polynomial degrees
abts_pol2 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 2),
              data = abts,
              family = binomial)
abts_pol3 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 3),
              data = abts,
              family = binomial)
abts_pol4 <- glm(formula = Inhibition_persentage ~ poly(Concentration, 4),
              data = abts,
              family = binomial)

# Compare models
AIC(abts_m, abts_pol2, abts_pol3, abts_pol4)

# Display degrees
tab_model(abts_pol4)

# Installing performance package
#install.packages("performance")
# Loading performance package
library(performance)

# Check model assumptions
check_model(abts_m)


abts_m1 <- mgcv::gam(Inhibition_persentage ~ Concentration,
                            abts, family = quasibinomial)
plot(abts_m1, all.terms = TRUE)


# Check all model assumptions visually
check_model(abts_m)


# Installing lme4 package
install.packages("lme4",
                 repos=c("http://lme4.r-forge.r-project.org/repos",
                         getOption("repos")[["CRAN"]]))
# Loading lme4 package
library(lme4)
# Mixed-effects logistic regression
antioxidant_m2 <- glmer(Inhibition_persentage ~ Location_factor +
                          Age_factor + Light_factor + Concentration +
                          Analysis_day,
                        abts,
                        quasibinomial)
check_model(abts_m)


# Installing ggeffects package
#install.packages("ggeffects")
# Loading ggeffects package
library(ggeffects)



# Extract predictions
#ggeffect(abts_m)

ggpredict(abts_m)

# Be careful
predict_response(abts_m, terms = "Concentration [all]")


# Visualize predictions
antioxidant_plot <- ggpredict(abts_m) %>%
  plot() %>%
  sjPlot::plot_grid()





```

Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

# Build the model
abts_m <- glm(formula = Inhibition_persentage ~ Location_factor + Age_factor +
                Light_factor + Analysis_day,
              data = abts,
              family = quasibinomial)
# Check model assumptions
check_model(abts_m)



```

We observed that Analysis day and location factor have collinearity. We inspect the data again with out location factor.

```{r pressure, echo=FALSE}

# Build the model
abts_no_location <- glm(formula = Inhibition_persentage ~ Age_factor +
                          Light_factor + Analysis_day,
                        data = abts,
                        family = quasibinomial)
# Check model assumptions
check_model(abts_no_location)


```

Here description of result.

```{r pressure, echo=FALSE}

# Extract prediction
ggeffect(abts_no_location)


```

Visualization the results by a plot.

```{r pressure, echo=FALSE}

# Library to use "%>%"
library(dplyr)
# Library to use "ggsave"
library(tidyverse)
# Visualize predictions
abts_plot <- ggeffect(abts_no_location) %>%
  plot()  %>%
  sjPlot::plot_grid()
# Safe result plot
ggsave('Result/Antioxidant_Result/ABTS_Result.pdf', plot = abts_plot,
       width = 8, height = 6, device='pdf', dpi="print")
ggsave('Result/Antioxidant_Result/ABTS_Result.png', plot = abts_plot,
       width = 8, height = 6, device='png', dpi="print")


```

Visualization the results by a table.

```{r pressure, echo=FALSE}

# Installing gtsummary package
#install.packages("gtsummary")

# Loading gtsummary package
library(gtsummary)

# Display contrasts / post-hocs / pairwise camparisons
abts_table <- tbl_regression(abts_no_location,
                             exponentiate = TRUE,
                             add_pairwise_contrasts = TRUE,
                             contrasts_adjust = "bonferroni",
                             pairwise_reverse = FALSE,
                             pvalue_fun = ~style_pvalue(.x, digits = 3)) %>%
  add_significance_stars(hide_p = F, hide_se = T, hide_ci = F) %>%
  bold_p()
abts_table



```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
