---
title: "Spectral deconvolution of *Ilex guayusa* GC-MS metabolomics"
author: "Jefferson Pastuna, Edison Gonzales"
date: "2024-06-09"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

# Introduction

The present document aims to record the GC-EI (Q)MS spectral deconvolution procedure of the *Ilex guayusa* volatile profile. A brief explanation, the code, and graphics are included for each step.

The workflow used is taken from the paper [eRah: A Computational Tool Integrating Spectral Deconvolution and Alignment with Quantification and Identification of Metabolites in GC/MS-Based Metabolomics](https://doi.org/10.1021/acs.analchem.6b02927). It offers a wide variety of functions to automatically detect and deconvolve the spectra of the compounds appearing in GC–MS chromatograms.

# Before to start

The "eRah" package accepts raw data files (netCDF or mzXML) obtained in GC–q/MS, GC-TOF/MS, and GC-qTOF/MS (using nominal mass) equipment.

In this case, the data was obtained with a GC-2030 gas chromatograph coupled to a GCMS-QP2020 NX quadrupole mass spectrometer operating in electron ionization (GC-EI-q/MS). The raw MS data (.qgd) were converted to (netCDF) format using the proprietary software of the instrument GCMS Postrun Analysis 4.53SP1. The data was organized in an experiment folder named Data_to_eRah. This contains three class-folders called ’Process_Blank’, ’Quality_Control’, and ’Samples’, each containing the sample files for that class.

# eRah package workflow

The eRah package is installed and loaded.

```{r echo=TRUE, message=FALSE, warning=FALSE}

# eRah package installation
#install.packages('erah')
# eRah library loading
library(erah)

```

Delete unwanted files in the experiment folder and create the instrumental and phenotype files.

```{r echo=TRUE}

# Delete all file that are not in folders
unlink('Data/Data_to_eRah/*')
# Data folder path
createdt('Data/Data_to_eRah/')

```

The previously created files (instrumental and phenotype) were loaded in R. Then, the MetaboSet object was created to store the sample metadata and processing results.

**NOTE.** The instrumental and phenotype files created by eRah were relocated to a new directory. In both files, the semicolon delimiter has been changed to a comma delimiter. The complete directory path of the chromatograms has been added to the instrumental file.

```{r echo=TRUE}

# Loading (*.mzXML) chromatograms mane (instrumental file)
#instrumental <- read.csv('Data/Metadata_to_eRah/Metadata_inst_mzXML.csv')
# If (*.mzXML) did not work
# Loading (*.CDF) chromatograms
instrumental <- read.csv('Data/Metadata_to_eRah/Metadata_inst_CDF.csv')
# Loading metadata of the chromatograms (phenotype file)
phenotype <- read.csv('Data/Metadata_to_eRah/Metadata_pheno.csv')
# Merge of metadata information with chromatograms
raw_data <- newExp(instrumental = instrumental,
                   phenotype = phenotype,
                   info = 'I. guayusa volatilome')

```

## Compound Deconvolution

To improve the accuracy and relevance of chemical analysis, parameters for composite deconvolution are specified, specific criteria on peak width, minimum height, and noise threshold are defined, and certain ranges of m/z values are excluded from processing.

```{r echo=TRUE}

dec_par <- setDecPar(min.peak.width = 0.7,
                     min.peak.height = 500,
                     noise.threshold = 50,
                     avoid.processing.mz = c(59:64,73:75,147:149),
                     analysis.time = c(3.1,50))

```

To carry out a process in parallel, we use the "future" package, which allows us to execute tasks in parallel, improving efficiency and processing speed simultaneously.

```{r echo=TRUE}

plan(future::multisession,
     workers = 14)

```

We proceed to the deconvolution of compounds using the parameters specified in "dec_par". The results obtained will be saved in (dec_data).

```{r echo=TRUE, warning=FALSE}

dec_data <- deconvolveComp(raw_data,
                           dec_par)

```

## Alignment

Parameters are defined for alignment and applied to the data.

```{r echo=TRUE}

# Alignment parameters
alig_par <- setAlPar(min.spectra.cor = 0.70,
                     max.time.dist = 5,
                     mz.range = 50:500)
# Alignment
peak_alig <- alignComp(dec_data,
                       alParameters = alig_par)

```

## Missing Compound Recovery

This step is used to recover missing compounds in the samples, the parameter only requires the number of minimum values for which a compound wants to be ’re-searched’ in the samples.

```{r echo=TRUE, warning=FALSE}

peak_find <- recMissComp(peak_alig,
                         min.samples = 3)

```

Exporting alignment feature list for blank subtraction with ["notame"](https://doi.org/10.3390/metabo10040135) R package. The procedures used are available in the "Multivariate_Statistics" notebook of GitHub repository (https://github.com/IKIAM-NPLab/I_guayusa_volatilome). Only high-quality features were used for metabolite identification.

```{r echo=TRUE}

# Extracting alignment feature list
feat_list <- alignList(peak_find,
                       by.area = FALSE)
# Exporting alignment feature list
#write.csv(feat_list,
#          file = "Result/eRah_Result/erah_Export_2Notame.csv")

```

# Identification

Metabolite identification was by comparing all the spectra found against a reference database. The eRah default database was used to metabolite identification. Metabolite identification was improve by exporting all the spectra found (.msp) in eRah to NIST MS Search 2.4 software.

```{r echo=TRUE, warning=FALSE}

# Identification
peak_iden <- identifyComp(peak_find,
                          id.database = mslib,
                          mz.range = NULL,
                          n.putative = 1)

```

Exporting spectra to NIST MS Search software identification with NIST-20 library.

```{r echo=TRUE}

#export2MSP(peak_iden,
#           store.path = "Result/eRah_Result",
#           alg.version = 2)

```

