---
title: "Effect of age and light on the volatile profile of *Ilex guayusa* leaves - Univariate statistical analysis"
author: "Jefferson Pastuña"
date: "2024-04-17"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

# Introduction

This R Script aims to record the procedure for analyzing the volatile profile of *Ilex guayusa* leaves under different age and light conditions. Each step has a brief explanation, as well as code and graphics.

The previous preprocessed data in the [multivariate statistics notebook](https://doi.org/10.3390/metabo10040135) will be used here.

# Volcano plot

The generalized logarithm (glog) transformed data and the PQN normalized data were used for volcano plot analysis. The two-sample t-test method (was calculated using glog transformed data) was used for the "y" axis of the volcano plot, and fold change (was calculated using PQN normalized data) was used for the "x" axis. We plotted two volcanoes, one with SubQC data and the other with individual data.

The first step was data extraction.

## SubQC

SubQC data extraction for t-test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Removal of sample group from the dataset
volc_glog <- glog_set[, glog_set$Analysis_group != "Sample"]
pData(volc_glog) <- droplevels(pData(volc_glog))
# Drop QC
volc_glog <- drop_qcs(volc_glog)
# Extracting SubQC light factor
volc_glog <- volc_glog[, volc_glog$Light_factor != "Other factor"]
pData(volc_glog) <- droplevels(pData(volc_glog))

```

SubQC data extraction for fold change calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Removal of sample group from the dataset
volc_pqn <- ppm_noflag[, ppm_noflag$Analysis_group != "Sample"]
pData(volc_pqn) <- droplevels(pData(volc_pqn))
# Drop QC
volc_pqn <- drop_qcs(volc_pqn)
# Extracting SubQC light factor
volc_pqn <- volc_pqn[, volc_pqn$Light_factor != "Other factor"]
pData(volc_pqn) <- droplevels(pData(volc_pqn))

```

Calculation of t-test and fold change.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Fold change between shade and light levels
volc_fc <- fold_change(volc_pqn, group = "Light_factor")
# two-sample t-test performing
volc_t <- perform_t_test(volc_glog, formula_char = "Feature ~ Light_factor")
# Adding the fold change to the t-test data
volc_data <- left_join(volc_t, volc_fc)
# Log-transform for visualization
volc_data$logP <- -log10(volc_data$Light_vs_Shade_t_test_P)
volc_data$log2_fc <- log2(volc_data$Shade_vs_Light_FC)

```

Volcano plot of SubQC.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.width = 9, fig.height = 15}

# Determine point colors based on significance and fold change
volc_data <- volc_data %>%
  mutate(point_color = case_when(
    Light_vs_Shade_t_test_P < 0.05 & log2_fc < -1 ~ "Down", # significantly down
    Light_vs_Shade_t_test_P < 0.05 & log2_fc > 1 ~ "Up",    # significantly up
    log2_fc < 1 & log2_fc > -1 ~ "No fold change",          # fold change < 2
    Light_vs_Shade_t_test_P > 0.05 & log2_fc < -1 | 
      Light_vs_Shade_t_test_P > 0.05 & log2_fc > 1 ~ "Fold change")) # fold change > 2
# Creating a new small table of the annotated compounds
## Keep metabolites with p-value < 0.05
volc_compouds <- subset(volc_data, Light_vs_Shade_t_test_P < 0.05 |
                             point_color == "Fold change")
volc_compouds <- left_join(meta_table, volc_compouds)
# Volcano plot
vc_plot <- ggplot(volc_data, aes(log2_fc, logP, color = point_color)) +
  geom_point(size = 2, alpha = 0.4) +
  scale_colour_manual(values = c("No fold change" = "#808080",
                                 "Down" = "#4F9D4EFF",
                                 "Up" = "#ff80ff",
                                 "Fold change" = "#F5BC5CFF")) +
  theme_classic() +
  #theme(legend.position = "none") +
  #geom_point(data = volc_compouds,
  #           aes(shape = meta_table$Identification_level,
  #               color = meta_table$Identification_level),
  #           size = 2) +
  labs(#shape = 'Identification level',
       color = 'Color key') +
  ggrepel::geom_label_repel(data = volc_compouds,
                            aes(label = meta_table$Metabolite),
                            color = "black",
                            box.padding = 0.37,
                            label.padding = 0.22,
                            label.r = 0.30,
                            cex = 2.5,
                            max.overlaps = 20,
                            min.segment.length = 0,
                            seed = 42,) +
  xlab(bquote(Log[2](Shade/Light))) + ylab(bquote(-Log[10](p-value))) +
  scale_y_continuous(limits = c(0,4), labels = function(i) 10^-i,
                     expand=c(0.003, 0.003)) +
  theme(legend.position = c(0.90, 0.947),
        legend.background = element_rect(fill = "white", color = "black")) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(fill= "transparent")) +
  #geom_vline(xintercept = 1, linetype = "longdash", colour="gray") +
  #geom_vline(xintercept = -1, linetype = "longdash", colour="gray") +
  geom_vline(xintercept = 0, linetype = "longdash", colour="gray") +
  geom_hline(yintercept = -log10(0.05), linetype = "longdash", colour="gray")
vc_plot
# Save plot
#ggsave('Result/notame_Result/figure_2b.pdf',
#       width = 5, height = 4, device='pdf', dpi="print")

```

## Sample

Sample data extraction for t-test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Removal of sample group from the dataset
sp_volc_glog <- glog_set[, glog_set$Analysis_group != "Sub_Quality_Control"]
pData(sp_volc_glog) <- droplevels(pData(sp_volc_glog))
# Drop QC
sp_volc_glog <- drop_qcs(sp_volc_glog)

```

Sample data extraction for fold change calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Removal of sample group from the dataset
sp_volc_pqn <- ppm_noflag[, ppm_noflag$Analysis_group != "Sub_Quality_Control"]
pData(sp_volc_pqn) <- droplevels(pData(sp_volc_pqn))
# Drop QC
sp_volc_pqn <- drop_qcs(sp_volc_pqn)

```

Calculation of t-test and fold change.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Fold change between shade and light levels
sp_volc_fc <- fold_change(sp_volc_pqn, group = "Light_factor")
# two-sample t-test performing
sp_volc_t <- perform_t_test(sp_volc_glog,
                            formula_char = "Feature ~ Light_factor")
# Adding the fold change to the t-test data
sp_volc_data <- left_join(sp_volc_t, sp_volc_fc)
# Log-transform for visualization
sp_volc_data$logP <- -log10(sp_volc_data$Light_vs_Shade_t_test_P)
sp_volc_data$log2_fc <- log2(sp_volc_data$Shade_vs_Light_FC)

```

Volcano plot of Sample.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.width = 9, fig.height = 15}

# Determine point colors based on significance and fold change
sp_volc_data <- sp_volc_data %>%
  mutate(point_color = case_when(
    Light_vs_Shade_t_test_P < 0.05 & log2_fc < -1 ~ "Down", # significantly down
    Light_vs_Shade_t_test_P < 0.05 & log2_fc > 1 ~ "Up",    # significantly up
    log2_fc < 1 & log2_fc > -1 ~ "No fold change",          # fold change < 2
    Light_vs_Shade_t_test_P > 0.05 & log2_fc < -1 | 
      Light_vs_Shade_t_test_P > 0.05 & log2_fc > 1 ~ "Fold change")) # fold change > 2
# Creating a new small table of the annotated compounds
## Keep metabolites with p-value < 0.05
sp_volc_compouds <- subset(sp_volc_data, Light_vs_Shade_t_test_P < 0.05 |
                             point_color == "Fold change")
sp_volc_compouds <- left_join(meta_table, sp_volc_compouds)
# Volcano plot
sp_vc_plot <- ggplot(sp_volc_data, aes(log2_fc, logP, color = point_color)) +
  geom_point(size = 2, alpha = 0.4) +
  scale_colour_manual(values = c("No fold change" = "#808080",
                                 "Down" = "#4F9D4EFF",
                                 "Up" = "#ff80ff",
                                 "Fold change" = "#F5BC5CFF")) +
  theme_classic() +
  #theme(legend.position = "none") +
  #geom_point(data = volc_compouds,
  #           aes(shape = meta_table$Identification_level,
  #               color = meta_table$Identification_level),
  #           size = 2) +
  labs(#shape = 'Identification level',
       color = 'Color key') +
  ggrepel::geom_label_repel(data = sp_volc_compouds,
                            aes(label = meta_table$Metabolite),
                            color = "black",
                            box.padding = 0.37,
                            label.padding = 0.22,
                            label.r = 0.30,
                            cex = 2.5,
                            max.overlaps = 20,
                            min.segment.length = 0,
                            seed = 42,) +
  xlab(bquote(Log[2](Shade/Light))) + ylab(bquote(-Log[10](p-value))) +
  scale_y_continuous(limits = c(0,4), labels = function(i) 10^-i,
                     expand=c(0.003, 0.003)) +
  theme(legend.position = c(0.90, 0.947),
        legend.background = element_rect(fill = "white", color = "black")) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(fill= "transparent")) +
  #geom_vline(xintercept = 1, linetype = "longdash", colour="gray") +
  #geom_vline(xintercept = -1, linetype = "longdash", colour="gray") +
  geom_vline(xintercept = 0, linetype = "longdash", colour="gray") +
  geom_hline(yintercept = -log10(0.05), linetype = "longdash", colour="gray")
sp_vc_plot
# Save plot
#ggsave('Result/notame_Result/figure_2b.pdf',
#       width = 5, height = 4, device='pdf', dpi="print")

```

# Tukey test

The tukey test was implemented to support the multivariate statistic by analyzing individual identified metabolites between age factors (with three levels: early, medium, and late). Only metabolites with data homoscedasticity were considered for the ANOVA test, and metabolites with significance (p-value < 0.05) were used for the Tukey test. The generalized logarithm (glog) transformed data was used for the Tukey test.

Previously, we tested with the auto-scaled data and found that seven metabolites did not satisfy the data homoscedasticity. On the other hand, when we tested with the glog-transformed data, only two metabolites did not satisfy the data homoscedasticity. For the above reasons, we used log-transformed data for the Tukey test.

We analyzed two batches of data, one of the SubQC data and the other of the individual data.

## SubQC

SubQC data extraction for Tukey test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Drop QC
tk_no_qc <- drop_qcs(glog_set)
# Drop samples
tk_no_sample <- tk_no_qc[, tk_no_qc$Analysis_group != "Sample"]
pData(tk_no_sample) <- droplevels(pData(tk_no_sample))
# Extracting levels of age factor
tk_age <- tk_no_sample[, tk_no_sample$Age_factor != "Other factor"]
pData(tk_age) <- droplevels(pData(tk_age))
# Extracting identified metabolites
tk_age_set <- tk_age[!is.na(tk_age@featureData@data$Metabolite),]

```

Preparing data and metadata for Tukey test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Extracting feature height table
tk_height <- exprs(tk_age_set)
# Extracting sample information
tk_pdata <- tk_age_set@phenoData@data
tk_pdata <- data.frame(Sample_ID = tk_pdata$Sample_ID,
                       Age_factor = tk_pdata$Age_factor)
# Extracting feature information
tk_fdata <- tk_age_set@featureData@data
tk_fdata <- data.frame(Feature_ID = tk_fdata$Feature_ID,
                       Metabolite = tk_fdata$Metabolite)
# Transposing feature height table
tk_height <- t(tk_height)
# Feature height table to dataframe
tk_height <- as.data.frame(tk_height)
# Convert the row names to sample ID
tk_height <- tk_height %>% 
  mutate(Sample_ID = rownames(tk_height))
# Adding age factor as variable
tk_height <- left_join(tk_pdata, tk_height)

```

The data is ready to test the data homoscedasticity. The Bartlett test was used to inspect the homogeneity of variance.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Create two empty vectors to add results
Feature_ID <- c()
bartlett_pValue <- c()
# Batch analysis
for(i in 3:80) {
  bar_res <- bartlett.test(tk_height[, i] ~ tk_height$Age_factor,
                           data = tk_height)
  bartlett_pValue <- c(bartlett_pValue, bar_res[["p.value"]])
  Feature_ID <- c(Feature_ID, names(tk_height[i]))
  tk_bar_res <- data.frame(Feature_ID, bartlett_pValue)
}
# Adding a description of the p-value
tk_bar_res <- tk_bar_res %>%
  mutate(bartlett_group = ifelse(bartlett_pValue > 0.05,
                                 "equal variance",
                                 "Unequal variance"))

```

Most metabolites showed equality of variances according to the Bartlett test. The next step will be the ANOVA test.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Create two empty vectors to add results
Feature_ID <- c()
anova_pValue <- c()
# Batch analysis
for(i in 3:80) {
  anova_model <- glm(tk_height[, i] ~ tk_height$Age_factor, data = tk_height)
  anova_res <- anova(anova_model, test = "LRT")
  anova_pValue <- c(anova_pValue, anova_res$`Pr(>Chi)`[2])
  Feature_ID <- c(Feature_ID, names(tk_height[i]))
  tk_anova_res <- data.frame(Feature_ID, anova_pValue)
}
# Adding a description of the p-value
tk_anova_res <- tk_anova_res %>%
  mutate(anova_group = ifelse(anova_pValue > 0.05,
                              "Not Significant",
                              "Significant"))

```

Consequently, the features with significance (p-value < 0.05) according to ANOVA were used for the Tukey test. Tukey test groups the different variables (in this case, the levels of age factor: early, medium, and late) using significance tests.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# agricolae package installation and library loadding
#install.packages("agricolae", repos = "https://cran.r-project.org")
library(agricolae)
# Create two empty vectors to add results
Feature_ID <- c()
tk_group_early <- c()
tk_group_late <- c()
tk_group_medium <- c()
# Batch analysis
for(i in 3:80) {
  tk_model <- aov(tk_height[, i] ~ tk_height$Age_factor, data = tk_height)
  tk_tes <- HSD.test(tk_model, "tk_height$Age_factor", group = TRUE)
  tk_group_early <- c(tk_group_early, tk_tes[["groups"]]["Early",2])
  tk_group_late <- c(tk_group_late, tk_tes[["groups"]]["Late",2])
  tk_group_medium <- c(tk_group_medium, tk_tes[["groups"]]["Medium",2])
  Feature_ID <- c(Feature_ID, names(tk_height[i]))
  tk_res <- data.frame(Feature_ID, tk_group_early, tk_group_late,
                       tk_group_medium)
  tk_res
}

```

Merge all results and add a flag in the Tukey result column: "-" for metabolites with Unequal variance according to the Bartlett test and "ns" for metabolites not significant according to the ANOVA test.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Merge all results obtained in the Tukey test
tk_result <- left_join(tk_bar_res, tk_anova_res)
tk_result <- left_join(tk_result, tk_res)
# Flag "-" metabolites that has unequal variance
for(i in 1:78) {
  if (tk_result$bartlett_group[i] == "Unequal variance") {
		tk_result$tk_group_early[i] <- "-"
		tk_result$tk_group_medium[i] <- "-"
		tk_result$tk_group_late[i] <- "-"
  }
  if (tk_result$anova_group[i] == "Not Significant" &
      tk_result$bartlett_group[i] == "equal variance") {
		tk_result$tk_group_early[i] <- "ns"
		tk_result$tk_group_medium[i] <- "ns"
		tk_result$tk_group_late[i] <- "ns"
}
 tk_result
}

```

Creating a Tukey result table to use in the heatmap ([multivariate statistics notebook](https://doi.org/10.3390/metabo10040135)).

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Tukey table to heatmap
hm_tk_table <- data.frame(Feature_ID = tk_result$Feature_ID,
                          Early = tk_result$tk_group_early,
                          Late = tk_result$tk_group_late,
                          Medium = tk_result$tk_group_medium)

```

Print Tukey results in a table.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# The "gt" package installation
#install.packages("gt")
library(gt)
# Adding metabolite name
tk_table <- left_join(tk_fdata, hm_tk_table)
# Deleting the Feature_ID column
tk_table <- subset(tk_table, select = -c(Feature_ID) )
tk_table %>%
  gt()%>% 
  tab_spanner(
    label = "Tukey test",
    columns = 2:4)%>%
  tab_source_note(source_note = "Note: Different lowercase letter within the same row indicates significant differences based on Tukey's multiple comparisons (p-value < 0.05, n = 3). '-' represents features with unequal variance according to Bartlett's test. 'ns' are the features that do not have statistical differences (p-value > 0.05) according to the ANOVA test.")

```

## Sample

Sample data extraction for Tukey test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Drop factor pooled (subQC)
tk_no_subqc <- tk_no_qc[, tk_no_qc$Analysis_group != "Sub_Quality_Control"]
pData(tk_no_subqc) <- droplevels(pData(tk_no_subqc))
# Extracting identified metabolites
tk_no_subqc <- tk_no_subqc[!is.na(tk_no_subqc@featureData@data$Metabolite),]

```

Preparing data and metadata for Tukey test calculation.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Extracting feature height table
tk_spl_height <- exprs(tk_no_subqc)
# Extracting sample information
tk_spl_pdata <- tk_no_subqc@phenoData@data
tk_spl_pdata <- data.frame(Sample_ID = tk_spl_pdata$Sample_ID,
                           Age_factor = tk_spl_pdata$Age_factor)
# Extracting feature information
tk_spl_fdata <- tk_no_subqc@featureData@data
tk_spl_fdata <- data.frame(Feature_ID = tk_spl_fdata$Feature_ID,
                           Metabolite = tk_spl_fdata$Metabolite)
# Transposing feature height table
tk_spl_height <- t(tk_spl_height)
# Feature height table to dataframe
tk_spl_height <- as.data.frame(tk_spl_height)
# Convert the row names to sample ID
tk_spl_height <- tk_spl_height %>% 
  mutate(Sample_ID = rownames(tk_spl_height))
# Adding age factor as variable
tk_spl_height <- left_join(tk_spl_pdata, tk_spl_height)

```

The data is ready to test the data homoscedasticity. The Bartlett test was used to inspect the homogeneity of variance.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Create two empty vectors to add results
Feature_ID <- c()
bartlett_pValue <- c()
# Batch analysis
for(i in 3:80) {
  bar_res <- bartlett.test(tk_spl_height[, i] ~ tk_spl_height$Age_factor,
                           data = tk_spl_height)
  bartlett_pValue <- c(bartlett_pValue, bar_res[["p.value"]])
  Feature_ID <- c(Feature_ID, names(tk_spl_height[i]))
  tk_spl_bar_res <- data.frame(Feature_ID, bartlett_pValue)
}
# Adding a description of the p-value
tk_spl_bar_res <- tk_spl_bar_res %>%
  mutate(bartlett_group = ifelse(bartlett_pValue > 0.05,
                                 "equal variance",
                                 "Unequal variance"))

```

Some metabolites showed equality of variances according to the Bartlett test. The next step will be the ANOVA test.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Create two empty vectors to add results
Feature_ID <- c()
anova_pValue <- c()
# Batch analysis
for(i in 3:80) {
  anova_model <- glm(tk_spl_height[, i] ~ tk_spl_height$Age_factor,
                     data = tk_spl_height)
  anova_res <- anova(anova_model, test = "LRT")
  anova_pValue <- c(anova_pValue, anova_res$`Pr(>Chi)`[2])
  Feature_ID <- c(Feature_ID, names(tk_spl_height[i]))
  tk_spl_anova_res <- data.frame(Feature_ID, anova_pValue)
}
# Adding a description of the p-value
tk_spl_anova_res <- tk_spl_anova_res %>%
  mutate(anova_group = ifelse(anova_pValue > 0.05,
                              "Not Significant",
                              "Significant"))

```

Consequently, the features with significance (p-value < 0.05) according to ANOVA were used for the Tukey test. Tukey test groups the different variables (in this case, the levels of age factor: early, medium, and late) using significance tests.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Create two empty vectors to add results
Feature_ID <- c()
tk_group_early <- c()
tk_group_late <- c()
tk_group_medium <- c()
# Batch analysis
for(i in 3:80) {
  tk_model <- aov(tk_spl_height[, i] ~ tk_spl_height$Age_factor,
                  data = tk_spl_height)
  tk_tes <- HSD.test(tk_model, "tk_spl_height$Age_factor", group = TRUE)
  tk_group_early <- c(tk_group_early, tk_tes[["groups"]]["Early",2])
  tk_group_late <- c(tk_group_late, tk_tes[["groups"]]["Late",2])
  tk_group_medium <- c(tk_group_medium, tk_tes[["groups"]]["Medium",2])
  Feature_ID <- c(Feature_ID, names(tk_spl_height[i]))
  tk_spl_res <- data.frame(Feature_ID, tk_group_early, tk_group_late,
                       tk_group_medium)
  tk_spl_res
}

```

Merge all results and add a flag in the Tukey result column: "-" for metabolites with Unequal variance according to the Bartlett test and "ns" for metabolites not significant according to the ANOVA test.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Merge all results obtained in the Tukey test
tk_spl_result <- left_join(tk_spl_bar_res, tk_spl_anova_res)
tk_spl_result <- left_join(tk_spl_result, tk_spl_res)
# Flag "-" metabolites that has unequal variance
for(i in 1:78) {
  if (tk_spl_result$bartlett_group[i] == "Unequal variance") {
		tk_spl_result$tk_group_early[i] <- "-"
		tk_spl_result$tk_group_medium[i] <- "-"
		tk_spl_result$tk_group_late[i] <- "-"
  }
  if (tk_spl_result$anova_group[i] == "Not Significant" &
      tk_spl_result$bartlett_group[i] == "equal variance") {
		tk_spl_result$tk_group_early[i] <- "ns"
		tk_spl_result$tk_group_medium[i] <- "ns"
		tk_spl_result$tk_group_late[i] <- "ns"
}
 tk_spl_result
}

```

Creating a Tukey result table to use in the heatmap ([multivariate statistics notebook](https://doi.org/10.3390/metabo10040135)).

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Tukey table to heatmap
hm_tk_table <- data.frame(Feature_ID = tk_group$Feature_ID,
                          Early = tk_group$tk_group_early,
                          Late = tk_group$tk_group_late,
                          Medium = tk_group$tk_group_medium)
# Replacing NAs by "ns" (not Significant)
hm_tk_table[is.na(hm_tk_table)]  <- "ns"

```












Finish a record.

```{r}

finish_log()

```



